### 2016-09-01 paper.tex
Add Vince edits

The mechanical power, total rotor drag, and near-wake velocity of a 1:6 scale
(1.075 m diameter) DOE/Sandia National Labs Reference Model 2 (RM2)
vertical-axis cross-flow turbine were measured experimentally in a towing tank,
to provide a comprehensive open dataset for validating numerical models.
Performance was measured for a range of tip speed ratios and at multiple
Reynolds numbers by varying the rotor's angular velocity and tow carriage speed,
respectively. A peak power coefficient $C_P = 0.37$ and rotor drag coefficient
$C_D = 0.84$ were observed at a tip speed ratio $\lambda_0 = 3.1$. A regime of
weak linear $Re$-dependence of the power coefficient was observed above a
turbine diameter Reynolds number $Re_D \approx 10^6$. The effects of support
strut drag on turbine performance were investigated by covering the rotor's NACA
0021 struts with cylinders. As expected, this modification drastically reduced
the rotor power coefficient. Strut drag losses were also measured for the NACA
0021 and cylindrical configurations with the rotor blades removed. For
$\lambda=\lambda_0$, wake velocity was measured at 1 m $(x/D=0.93)$ downstream.
Mean velocity, turbulence kinetic energy, and mean kinetic energy transport were
compared with results from a high solidity turbine acquired with the same test
apparatus. Like the high solidity case, mean vertical advection was calculated
to be the largest contributor to near-wake recovery. However, overall, lower
levels of streamwise wake recovery were calculated for the RM2 case---a
consequence of both the relatively low solidity and tapered blades reducing
blade tip vortex shedding---responsible for mean vertical advection---and lower
levels of turbulence caused by higher operating tip speed ratio and therefore
reduced dynamic stall. Datasets, code for processing and visualization, and a
CAD model of the turbine have been made publicly available
\cite{Bachant2016-RM2-data,Bachant2015-RM2-CAD}.

>>>>

The mechanical power, total rotor drag, and near-wake velocity of a 1:6 scale
model (1.075 m diameter) US Department of Energy's Reference Model vertical-axis
cross-flow turbine were measured experimentally in a towing tank, to provide a
comprehensive open dataset for validating numerical models. Performance was
measured for a range of tip speed ratios and at multiple Reynolds numbers by
varying the rotor's angular velocity and tow carriage speed, respectively. A
peak power coefficient $C_P = 0.37$ and rotor drag coefficient $C_D = 0.84$ were
observed at a tip speed ratio $\lambda_0 = 3.1$. A regime of weak linear
$Re$-dependence of the power coefficient was observed above a turbine diameter
Reynolds number $Re_D \approx 10^6$. The effects of support strut drag on
turbine performance were investigated by covering the rotor's NACA 0021 struts
with cylinders. As expected, this modification drastically reduced the rotor
power coefficient. Strut drag losses were also measured for the NACA 0021 and
cylindrical configurations with the rotor blades removed. For
$\lambda=\lambda_0$, wake velocity was measured at 1 m $(x/D=0.93)$ downstream.
Mean velocity, turbulence kinetic energy, and mean kinetic energy transport were
compared with results from a high solidity turbine acquired with the same test
apparatus. Like the high solidity case, mean vertical advection was calculated
to be the largest contributor to near-wake recovery. However, overall, lower
levels of streamwise wake recovery were calculated for the RM2 case---a
consequence of both the relatively low solidity and tapered blades reducing
blade tip vortex shedding---responsible for mean vertical advection---and lower
levels of turbulence caused by higher operating tip speed ratio and therefore
reduced dynamic stall. Datasets, code for processing and visualization, and a
CAD model of the turbine have been made publicly available
\cite{Bachant2016-RM2-data,Bachant2015-RM2-CAD}.
